% File: IGD_Core_Equations_v2.2.tex
\documentclass[12pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[english,ukrainian]{babel}

\title{Information--Geometric Dynamics (IGD)\\ \large Core Equations v2.2 (Consolidated)}
\author{Ordo Vacui \& Colleagues}
\date{\today}

\begin{document}
\maketitle

\section*{Core Structures}

Information manifold $\mathcal{M}_\theta$ endowed with:
\begin{itemize}
  \item $g_{ij}(\theta)$: Riemannian metric (Bures).
  \item $\mathcal{J}$: Complex structure (from A0).
  \item $J(\theta)$: Poisson operator (antisymmetric).
  \item $R(\theta)$: Dissipative operator (symmetric, positive semidefinite).
  \item $E(\theta)$ (Energy functional), $S(\theta)$ (Entropy functional).
\end{itemize}

\section*{IGD Equation (GENERIC Form)}

The evolution of the macrostate $\theta(\tau)$ is:
\begin{equation}
  \dot\theta = J(\theta)\,\nabla E(\theta) + R(\theta)\,\nabla S(\theta).
  \label{eq:IGD}
\end{equation}

\subsection*{Constraints and Conditions}

\begin{enumerate}
  \item \textbf{Degeneracy (Conservation Laws):}
    \begin{equation}
      J(\theta)\,\nabla S(\theta) = 0, \qquad
      R(\theta)\,\nabla E(\theta) = 0.
    \end{equation}

  \item \textbf{Entropy Production (Arrow of time):}
    \begin{equation}
      \frac{dS}{d\tau} = \nabla S^\top R(\theta)\,\nabla S \;\ge\; 0.
    \end{equation}

  \item \textbf{A10: Geometric (Kähler) Compatibility:}
    The quadruple $(g, \mathcal{J}, J, R)$ forms a compatible geometric structure. The metric and complex structure satisfy $g(\mathcal{J}u,\mathcal{J}v) = g(u,v)$. The symplectic form $\omega$ (associated with $J$) is given by:
    \begin{equation}
      \omega(u,v) = g(\mathcal{J}u,v).
    \end{equation}
    This ensures that the $J$-flow (reversible dynamics) is isometric. (Compatibility constraints on $R$ are an active area of research).

  \item \textbf{A9: Informational Locality:}
    The operators $J$ and $R$ must be quasi-local with respect to the informational adjacency graph defined by Mutual Information $I(i:j)$ between subsystems. This implies finite propagation speeds (Lieb-Robinson bounds).
\end{enumerate}

\section*{Emergent Time}

Information line element (Bures metric): $d\ell^2 = g_{ij}(\theta)\, d\theta^i d\theta^j$.
Physical time $t$ is parametrized (up to gauge $\lambda>0$):
\begin{equation}
  dt = \lambda(\theta)\, d\ell.
\end{equation}

\section*{Quantum Realization (Schematic)}

For a quantum system $(\rho, H)$: $E(\rho)=\mathrm{Tr}(\rho H)$, $S(\rho)=-\mathrm{Tr}(\rho\log\rho)$.
The IGD equation \eqref{eq:IGD}, constrained by A10 and A9, must reproduce generalized quantum dynamics.

\begin{align}
  \dot\rho_{\mathrm{rev}} &= -\frac{i}{\hbar}[H,\rho] && (\text{from } J\nabla E, \text{ constrained by A10})\\
  \dot\rho_{\mathrm{diss}} &= \mathcal{L}_{\mathrm{diss}}(\rho) && (\text{from } R\nabla S, \text{ e.g., Lindblad form})
\end{align}

\end{document}
