% File: IGD_SpinChain_Project_v1.1_Plan.tex
\documentclass[12pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[english,ukrainian]{babel}

\title{Project IGD--SpinChain v1.1: Concrete Plan}
\author{Ordo Vacui \& Colleagues}
\date{\today}

\begin{document}
\maketitle

\section*{Goal}

To demonstrate emergent geometry (A9) and the Area Law (T1 prototype) within the IGD framework using a solvable quantum spin chain model, realizing the axioms A9 and A10 explicitly.

\section{Model Definition: Transverse Field Ising Model (TFIM)}

We select the 1D TFIM on $N$ qubits ($\mathcal{H} = \bigotimes_{i=1}^N \mathbb{C}^2_i$):
\begin{equation}
  H = - J \sum_{i=1}^{N-1} \sigma^z_i \sigma^z_{i+1}
      - h \sum_{i=1}^N \sigma^x_i.
\end{equation}

\subsection*{Regime and Parameters}

We choose the \textbf{gapped phase} to ensure Area Law (T1) and locality (A9).
\begin{itemize}
  \item \textbf{Phase:} Ferromagnetic ($h < J$).
  \item \textbf{Parameters:} $J=1$ (energy scale), $h=0.5$. (We avoid the critical point $h=1$).
\end{itemize}

\section{Focus Manifolds ($\mathcal{M}_\theta$)}

We will utilize the exact solvability of the model via two complementary approaches:

\begin{description}
  \item[Focus 1: Gaussian Fermionic States (Analytical)]
    \begin{itemize}
      \item \textbf{Idea:} TFIM maps exactly to free fermions via Jordan-Wigner transformation. The ground state is a Gaussian state (Bogoliubov vacuum).
      \item \textbf{Parameters $\theta$:} The fermionic covariance matrix.
      \item \textbf{Advantage:} Allows for exact analytical calculation of information measures and the Bures metric.
    \end{itemize}

  \item[Focus 2: Matrix Product States (MPS) (Numerical/Variational)]
    \begin{itemize}
      \item \textbf{Idea:} MPS is the natural ansatz for 1D gapped systems and the basis for DMRG.
      \item \textbf{Parameters $\theta$:} Elements of the MPS tensors with bounded bond dimension.
      \item \textbf{Advantage:} Numerical verification and generalizability.
    \end{itemize}
\end{description}

\section{Key Observables (IGD Realization)}

\begin{itemize}
  \item \textbf{A9 (Locality):} Information graph weights $w_{ij} \propto I(i:j)_\rho$. We must show exponential decay $I(i:j) \sim e^{-|i-j|/\xi}$.
  \item \textbf{T1 (Area Law):} Entanglement entropy of a block $S(\ell)$. We must show saturation $S(\ell) \to \text{const}$ for $\ell \gg \xi$.
  \item \textbf{A10 (Geometry):} Bures metric $g^{\mathrm{B}}$ on the chosen focus manifold $\mathcal{M}_\theta$.
\end{itemize}

\section{Work Plan and Task Division}

\subsection*{Task 1: Analytics (Lead: Colleague)}

Focus 1 (Gaussian States) will be used for analytical derivation.

\begin{enumerate}
  \item \textbf{Fermionization and Diagonalization:} Perform Jordan-Wigner transformation and Bogoliubov diagonalization to find the ground state covariance matrix.
  \item \textbf{A9 Verification:} Analytically calculate the mutual information $I(i:j)$ and confirm exponential decay.
  \item \textbf{T1 Verification:} Analytically calculate the block entropy $S(\ell)$ and confirm the Area Law saturation.
  \item \textbf{A10 Realization:} Analytically derive the Bures metric on the manifold of Gaussian ground states. Determine the explicit form of the emergent 1D geometry.
\end{enumerate}

\subsection*{Task 2: Numerics (Lead: Ordo Vacui)}

Focus 2 (MPS/DMRG) will be used for numerical verification.

\begin{enumerate}
  \item \textbf{Ground State Calculation:} Use DMRG/MPS algorithms to find the ground state for finite $N$ chains at $h=0.5$.
  \item \textbf{Verification:} Numerically verify the analytical predictions for $I(i:j)$ and $S(\ell)$.
  \item \textbf{Visualization:} Visualize the informational graph $w_{ij}$.
\end{enumerate}

\subsection*{Outcome}

A complete "Proof of Concept" for the IGD framework, demonstrating the emergence of geometry and holographic properties from information principles in a concrete, solvable model.

\end{document}